---
title: "Fraud Detection Report"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(caret)
library(cluster)
library(factoextra)
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This report details the process and results of detecting potential fraudulent transactions using a bank transactions dataset. The primary methods include exploratory data analysis, data cleaning, and clustering via the K-Means algorithm. Results and insights are presented with visuals and summaries suitable.

# Data Exploration

## Load and Inspect the Data

```{r load-data}
data <- read.csv("C:/Users/arthu/Downloads/bank_transactions_data_2.csv")
summary(data)
```

## Check for Missing Values

```{r missing-values}
cat("\nMissing values:")
colSums(is.na(data))
```

## Visualize Key Variables

### Transaction Amount Distribution

```{r plot-transaction}
ggplot(data, aes(x = TransactionAmount)) + 
  geom_histogram(bins = 30, fill = "blue", color = "black") + 
  labs(title = "Distribution of Transaction Amount")
```

### Account Balance Distribution

```{r plot-balance}
ggplot(data, aes(x = AccountBalance)) + 
  geom_histogram(bins = 30, fill = "green", color = "black") + 
  labs(title = "Distribution of Account Balance")
```

# Fraud Detection Using K-Means Clustering

## Normalize the Data

```{r normalize}
numeric_scaled <- scale(data |> select(where(is.numeric)))
```

## Perform K-Means Clustering

```{r kmeans-clustering}
set.seed(42)
kmeans_result <- kmeans(numeric_scaled, centers = 5, nstart = 25)
data$KMeans_Cluster <- kmeans_result$cluster
```

## Identify Fraudulent Transactions

```{r fraud-detection}
centroids <- kmeans_result$centers
distances <- apply(numeric_scaled, 1, function(x) min(sqrt(colSums((t(centroids) - x)^2))))
data$KMeans_Distance <- distances

threshold <- mean(distances) + 3 * sd(distances)
data$KMeans_Fraud <- distances > threshold

fraud_points <- data |> filter(KMeans_Fraud == TRUE)
```

## Fraud Summary

```{r fraud-summary}
fraud_summary <- table(data$KMeans_Fraud)
cat("Fraud Summary:", fraud_summary)
```

# Results and Visualization

## Clustering Results

```{r cluster-plot}
ggplot(data, aes(x = TransactionAmount, y = AccountBalance, color = as.factor(KMeans_Cluster))) +
  geom_point(alpha = 0.7) +
  labs(title = "K-Means Clusters", x = "Transaction Amount", y = "Account Balance") +
  scale_color_viridis_d() +
  theme_minimal()
```

## Fraud Detection Visualization

```{r fraud-plot}
ggplot(data, aes(x = TransactionAmount, y = AccountBalance, color = as.factor(KMeans_Cluster))) +
  geom_point(alpha = 0.7) +
  geom_point(data = fraud_points, aes(x = TransactionAmount, y = AccountBalance), color = "red", size = 3) +
  labs(title = "K-Means Clusters with Fraud Points", x = "Transaction Amount", y = "Account Balance") +
  scale_color_viridis_d() +
  theme_minimal()
```

# Conclusion

This analysis identified a total of `r nrow(fraud_points)` potentially fraudulent transactions using the K-Means clustering algorithm. These outliers were flagged based on their distances from cluster centroids, exceeding the defined threshold.